---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", fig.path = "man/figures/README-",
  message = FALSE, warning = FALSE, error = FALSE, tidy = TRUE
)
```

# alfresco

[![Travis build status](https://travis-ci.org/leonawicz/alfresco.svg?branch=master)](https://travis-ci.org/leonawicz/alfresco)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/leonawicz/alfresco?branch=master&svg=true)](https://ci.appveyor.com/project/leonawicz/alfresco)
[![Coverage status](https://codecov.io/gh/leonawicz/alfresco/branch/master/graph/badge.svg)](https://codecov.io/github/leonawicz/alfresco?branch=master)

R is used to manipulate and analyze data that both feed into and come out of the ALFRESCO wildfire model. 
The `alfresco` R package contains functions for working with ALFRESCO input and output data. 
It is a core functional package in the SNAPverse collection of R packages.

## Features

The `alfresco` package includes functions that assist with the following analyses and processing pipelines:

* Fire events processing immediately following ALFRESCO runs
* Post-ALFRESCO run, automated Shiny app creation.
* Hierarchical data extractions from ALFRESCO run output geotiff map sets.
* Calculations for burn area ratios across multiple ALFRESCO run scenarios.
* Copying and organizing subsets of ALFRESCO outputs for consecutive model runs and in preparation for other analyses.
* Generating fire management options ratios geotiffs and related graphics.

This is a subset of eventual functionality. The package is under active development.

## Installation

You can install alfresco from github with:

```{r gh-instaallation, eval = FALSE}
# install.packages("devtoools")
devtools::install_github("leonawicz/alfresco")
```

```{r is_md, echo = FALSE}
is_md <- knitr::opts_knit$get("rmarkdown.pandoc.to") == "markdown_github-ascii_identifiers"
```

```{r md_doc3, echo = FALSE, results = "asis", eval = is_md}
cat('## Reference\n\n[Package reference and function documentation](https://leonawicz.github.io/alfresco/)')
```
